\Large \bf DRAFT: Direct-style process creation on Linux

### 2020-05-22 research/paper/paper.tex
paper

For \texttt{execve}, \verb|close_cloexec| is, in some sense, done automatically for us by the kernel;
except we don't get the chance to preserve live file descriptors handles,
without first unsetting \texttt{CLOEXEC} on them.
Again an extension to the kernel would be useful:
A new argument to \texttt{execve} which allows specifying a list of file descriptors to inherit
despite the \texttt{CLOEXEC} flag.
This would allow programs to avoid the pattern
of unsetting \texttt{CLOEXEC}, calling \texttt{execve}, and then immediately re-setting \texttt{CLOEXEC},
just to inherit a file descriptor across a single exec.
This might seem minor,
but unsetting \texttt{CLOEXEC} before \texttt{execve} is not possible
when the process calling \texttt{execve} has a shared file descriptor table,
because some other process might call \texttt{execve} as well,
and erroneously inherit the file descriptor.
\section{Evaluation}\label{evaluation}
Our goal is a powerful approach to process creation that is easy to use correctly.
Performance of our implementation is secondary,
but the overhead of our implementation must be low
relative to alternative approaches.

We evaluate the degree to which we meet this goal by answering the following questions:
\begin{compactitem}
\item
  Does direct-style process creation on Linux allow for simpler programs than the alternatives?
  We compare implementations of similar functionality using both direct-style and fork-style in \ref{ease},
  and find substantial benefits for direct-style.
\item
  Does our \texttt{rsyscall}-based Python direct-style process creation interface
  have acceptable performance relative to the Python standard library's spawn-style \texttt{subprocess.run}?
  We benchmark direct-style process creation against spawn-style in \ref{subprocess_bench},
  and find an acceptable level of overhead.
\item
  Does direct-style process creation perform well in the ``real world''?
  We discuss our positive experience with using direct-style process creation at \twosigma in \ref{realworld}.
\end{compactitem}
\subsection{Ease of programming with direct-style}\label{ease}
\begin{table}
\input{ease.tex}
\caption{Line counts with direct-style vs fork-style}
\label{tab:ease}
\end{table}
Does direct-style process creation on Linux allow for simpler programs than the alternatives?
To answer this, we compare the direct-style examples shown in section \ref{examples}
to programs with the same behavior implemented in fork-style.
Most of the examples cannot be implemented with typical Linux spawn-style interfaces such as \verb|posix_spawn|,
so we compare only to fork-style.

@@

Direct-style does not require any extra work to report errors;
errors are reported just like any other system call.
\subsection{Microbenchmark results}\label{subprocess_bench}
\begin{figure}[h!]
\centering
 \includegraphics[width=0.5\textwidth]{subprocess_bench}
 \caption{Python spawn-style vs direct-style performance under varying memory usage}
 \label{fig:subprocess_bench}
\end{figure}
Does our \texttt{rsyscall}-based Python direct-style \texttt{clone}
have acceptable performance relative to the Python standard library's spawn-style \texttt{subprocess.run}?

@@

The direct-style implementation stays around 2.2 milliseconds regardless of memory usage.
\texttt{subprocess.run}, which uses \texttt{fork},
starts at 1.4 milliseconds,
and then scales linearly with the amount of memory mapped, as expected,
up to 48.5 milliseconds at 4 GBs of mapped memory.
The cross-over point for this micro-benchmark is about 40 megabytes of memory mapped.

We perform another microbenchmark to evaluate the overhead of modifying the child process.
We simply run \texttt{getpid} in the child process on the same benchmark setup;
the average time per \texttt{getpid} call is .561 milliseconds.
The \texttt{getpid} syscall returns essentially instantly,
so we can regard this time as the amount of overhead incurred
by performing our child process setup system calls through \texttt{rsyscall}.
Through profiling, we've found that most of this per-syscall overhead is spent in Python,
rather than waiting for syscall responses from the child task.

@@

All this has allowed us to run our systems in a much more dynamic way than before.
We can freely start up arbitrary subsets of our systems on-demand,
spread across one or more hosts,
for development, testing, or production,
regardless of the configuration of that host,
without worrying about interference between instances,
and without a dependency on any external privileged services,
such as container runtimes.
This has become a key part of our development workflow.
\section{Related work}\label{related_work}
\subsection{Direct-style process creation}
Baumann 2019\cite{forkroad} briefly suggests cross-process operations as a possible replacement for fork on Unix,
positing many of the same advantages we find in practice.
Our work, though developed independently and largely completed before that paper,
is in many ways an elaboration on and implementation of that idea,
to which we owe a great debt.

@@

\bibliographystyle{plain}
% \bibliography{bibliography}
\end{document}

>>>>

For \texttt{execve}, \verb|close_cloexec| is, in some sense, done automatically for us by the kernel;
except we don't get the chance to preserve live file descriptors handles,
without first unsetting \texttt{CLOEXEC} on them.
Again an extension to the kernel would be useful:
A new argument to \texttt{execve} which allows specifying a list of file descriptors to inherit
despite the \texttt{CLOEXEC} flag.
This would allow programs to avoid the pattern
of unsetting \texttt{CLOEXEC}, calling \texttt{execve}, and then immediately re-setting \texttt{CLOEXEC},
just to inherit a file descriptor across a single exec.
This might seem minor,
but unsetting \texttt{CLOEXEC} before \texttt{execve} is not possible
when the process calling \texttt{execve} has a shared file descriptor table,
because some other process might call \texttt{execve} as well,
and erroneously inherit the file descriptor.
\section{Evaluation}\label{evaluation}
Our goal is a powerful approach to process creation that is easy to use correctly.
Performance of our implementation is secondary,
but the overhead of our implementation relative to alternative approaches
must be low.

We evaluate the degree to which we meet this goal by answering the following questions:
\begin{compactitem}
\item
  Does direct-style process creation on Linux allow for simpler programs than the alternatives?
  We compare implementations of similar functionality using both direct-style and fork-style in \ref{ease},
  and find substantial benefits for direct-style.
\item
  In Python, does our \texttt{rsyscall}-based direct-style process creation interface
  have acceptable performance relative to the standard library's spawn-style \texttt{subprocess.run}?
  We benchmark direct-style process creation against spawn-style in \ref{subprocess_bench},
  and find an acceptable level of overhead.
\item
  Does direct-style process creation perform well in the ``real world''?
  We discuss our positive experience with using direct-style process creation at \twosigma in \ref{realworld}.
\end{compactitem}
\subsection{Ease of programming with direct-style}\label{ease}
\begin{table}
\input{ease.tex}
\caption{Line counts with direct-style vs fork-style}
\label{tab:ease}
\end{table}
Does direct-style process creation on Linux allow for simpler programs than the alternatives?
To answer this, we compare the direct-style examples shown in section \ref{examples}
to programs with the same behavior implemented in fork-style.

Most of the examples cannot be implemented with typical Linux spawn-style interfaces such as \verb|posix_spawn|,
so we compare only to fork-style.

@@

Direct-style does not require any extra work to report errors;
errors are reported just like any other system call.
This is the main immediately-visible simplification of using direct-style process creation.

Direct-style also removes the need to be concerned about the state of the calling process,
though this simplification is not immediately visible in a small example.
The fork-style examples may break if there are multiple threads in the calling process,
and may be slow if the calling process has large amounts of memory mapped, as we'll see in section \ref{subprocess_bench}.
The direct-style examples do not have these limitations.
\subsection{Microbenchmark results}\label{subprocess_bench}
\begin{figure}[h!]
\centering
 \includegraphics[width=0.5\textwidth]{subprocess_bench}
 \caption{Python spawn-style vs direct-style performance under varying memory usage}
 \label{fig:subprocess_bench}
\end{figure}
In Python, does our \texttt{rsyscall}-based direct-style process creation interface
have acceptable performance relative to the standard library's spawn-style \texttt{subprocess.run}?

@@

The direct-style implementation stays around 2.2 milliseconds regardless of memory usage.
\texttt{subprocess.run}, which is implemented with \texttt{fork},
starts at 1.4 milliseconds,
and then scales linearly with the amount of memory mapped, as expected,
up to 48.5 milliseconds at 4 GBs of mapped memory.
The cross-over point for this micro-benchmark is about 40 megabytes of memory mapped.

We perform another microbenchmark to evaluate the overhead of modifying the child process.
We simply call \texttt{getpid} from Python
in both the child process and the parent process on the same benchmark setup.
The average time per \texttt{getpid} call is
3 microseconds when called in the parent process without going through \texttt{rsyscall},
and 561 miroseconds when called through \texttt{rsyscall} in the child process.
The difference, 558 microseconds,
is the amount of overhead incurred
by performing our child process setup system calls through \texttt{rsyscall}.
Through profiling, we've found that most of this per-syscall overhead is spent in Python,
rather than waiting for syscall responses from the child task.

@@

All this has allowed us to run our systems in a much more dynamic way than before.
We can freely start up arbitrary subsets of our systems on-demand,
spread across one or more hosts,
for development, testing, or production,
regardless of the configuration of that host,
without worrying about interference between instances,
and without a dependency on any external privileged services,
such as container runtimes.
This has become a key part of our development workflow.
\section{Related work}\label{related_work}
\subsection{Direct-style process creation}
Baumann 2019\cite{forkroad} briefly suggests cross-process operations as a possible replacement for fork on Unix,
positing many of the same advantages we find in practice.
Our work, though developed independently,
is in many ways an elaboration on and implementation of that idea,
to which we owe a great debt.

@@

\bibliographystyle{plain}
\bibliography{bibliography}
\end{document}
